Light inextensible strings under tension in the Schwarzschild geometry

### 2019-02-05 string_ajp.tex
more work for AJP submission

Integrating along a path between two
points~$p_1=\left(r_1,\phi_1\right)$ and~$p_2=\left(r_2,\phi_2\right)$
gives the {\em proper} length of the path, which is the length of a
string joining~$p_1$ and~$p_2$.  A taut but light string from~$p_1$
to~$p_2$ will adopt an extremal-length path.  Such paths may be found
by the calculus of variations; parametrizing a curve in terms
of~$r=r\left(\phi\right)$ gives us a path length of

@@

This differential equation does not appear to have a simple analytical
solution; solutions correspond to a taut string confined to Flamm's
paraboloid~\cite{flamm1916}.  It is straightforward to show that the
string has a point of inflection
if~$\left(r'\right)^2=2r\left(r-1\right)$.

>>>>

Elementary methods show that the proper length $\ell$ of radial strings has
some peculiar features.  Firstly, consider the proper length from
$r=1$ to $r=1+w$ where $w>0, \left|w\right|\ll 1$.  Then

\begin{equation}
  \ell=\int_{r=1}^{r=1+w}ds
  =2\sqrt{w} + \frac{1}{3}w^{3/2} + O(w^{5/2})
\end{equation}

This would indicate that, close to the Schwarzschild radius $r=1$,
large (but finite) amounts of string are required to connect two
points with slightly differing radial coordinates.  Passing to the
case $w\gg 1$ we have

\begin{equation}
  \ell  =\frac{1}{2} + \log(2) + w + \frac{1}{8}w^{-1} + O(w^{-2}),
\end{equation}

\noindent suggesting that there is ``extra radial space'', of length
over $1.19$ Schwarzschild radii, sensible at large distances from the
event horizon.  This might suggest that radial strings have to be
longer than expected.

\subsection{Nonradial string}
Integrating along a path between two
points~$p_1=\left(r_1,\phi_1\right)$ and~$p_2=\left(r_2,\phi_2\right)$
gives the {\em proper} length of the path, which is the length of an
inextensible string joining~$p_1$ and~$p_2$.  A taut but light string
from~$p_1$ to~$p_2$ will adopt an extremal-length path.  Such paths
may be found by the calculus of variations; parametrizing a curve in
terms of~$r=r\left(\phi\right)$ gives us a path length of

@@

This differential equation does not appear to be soluble analytically;
solutions correspond to a taut string confined to Flamm's
paraboloid~\cite{flamm1916}; see figure~\ref{flamm}.  It is
straightforward to show that the string has a point of inflection
if~$\left(r'\right)^2=2r\left(r-1\right)$.

\begin{figure}[h!] % Created by "flamm.R"
\centering
\includegraphics[width=70mm]{flamm_string.pdf}
\caption{Perspective fiew of Flamm's paraboloid with superimposed minimal-length path corresponding to a taut, light string}
\label{flamm}
\end{figure}
